---
title: "a3_markdown"
author: "Julien Hovan"
date: "2022-11-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

SETUP:
```{r}
library(readxl)
salary <- read_excel("/home/julienh/Downloads/salary2.xlsx")
x1 <- salary$SPENDING/1000
y <- salary$SALARY
fit.salary <- lm(y~x1)
summary(fit.salary)

```

Q1:
A)
```{r}
### A) Write R code to verify the calculation of the entries in the estimate column
n <- length(x1)
k <- 1

#transform x1 into a matrix with a column of ones (constant)
X <- matrix(x1)
X <- cbind(1, X)

# Estimates of Intercept (Beta_0) and x1 (Beta_1)
# Solving using OLS matrix algebra
# beta.hat = (X'X)^-1 X'y -> use solve() to get inverse and t() to transpose
beta.hat <- solve(t(X)%*%X) %*% t(X)%*%y
print("Beta hat values:")
print(beta.hat)
## -> values match the summary lm model

```
Answer: We can see that the values estimated using OLS matrix algebra match the beta estimates from the summary(lm) model



Q1B)
```{r}
### B) Verify numerically the orthogonality results concerning the residuals
# Compute the fit y_hat = X * B_hat
y.hat <- X %*% beta.hat
# Compute the residuals 
e.hat <- y - y.hat

# sum of residuals 
sum(e.hat)

# sum of (e.hat(xi - x_bar)) 
x_bar <- mean(x1)
sum(e.hat * (x1 - x_bar))

# sum of (e.hat * y.hat)
sum(e.hat * y.hat)

```
We can verify the results concerining the residuals, the justification for not being exactly 0 is the sample size of the data point is too small. As shown in class as n -> inf, both the residuals will go to 0 (ie. it is asympotomatically unbiased and goes to 0)

Q1C)
```{r}
### C) Compute the value of the entry in the Std. Error using entires given in the table then using the data directly
sig2 <- (t(e.hat)%*%e.hat) / (n-(k+1))

# Estimate of V[hat{beta}]
vcov_beta_hat <- c(sig2) * solve(t(X) %*% X)

# Estimate of standard errors
sqrt(diag(vcov_beta_hat))
cbind(beta.hat, sqrt(diag(vcov_beta_hat)))
```
We see that using the data directly, the value of the standard error for beta 0 and beta 1 hat.

Q1D)
```{r}

#### D) Write R code to compute the value of the ommitted entry for the residual standard erorr on line 15

# calculate sum of squared residuals
SSE <- sum(e.hat**2)
RSE <- sqrt(SSE/ n - (1+k))
RSE

```
Using the formula Residual_Standard_Error = SSE / n - (1+k)
RSE = 2278.739

Q1E)
The null hypothesis:
H0: B1 = 0
Alternate hypothesis:
H1: B1 != 0
Failing to reject the null implies that there is no linear relationship between X1 and Y since
Y = 0 * X1 + B0 + e = B0 + e

Decision Rule:
We reject H0 if the 100(1-alpha) percent confidence interval CI(B1) does not cover 0. The corresponding test statistic is 
T1 = beta1.hat - 0 / ese(beta1.hat)
```{r}
#### e) Test whether there is a linear association between SALARY and SPENDING using alpha = .05 -> state alternative hypothesis, decision rule, and conclsuion, what is the p-value of the test?
alpha = .005

ese.beta <- sqrt(diag(vcov_beta_hat))

test_statistic <- beta.hat / ese.beta
print("The t-statistic for beta1.hat is:")
test_statistic[2]

p.value <- 2*pt(-abs(test_statistic[2]), df= n -1)
print("The p value is: ")
p.value

```
Conclusion:
p-value < alpha = .005 to a significant degree,
-> There is sufficient evidence to conlucde that there is a significant linear relationship between x (SALARY) and y (SPENDING) because the correlation coefficient is significantly different from zero


Q2:
```{r}

```


Q3:
A) Fit simple linear regression model with g1 as the predictor -> second column of genetic data
```{r}
genetic_data <- read_excel("/home/julienh/Downloads/genetic_data2.xlsx")

genetic.lm.gene1 <- lm(genetic_data$y ~ genetic_data$X14046)

# Use t-test to compute the p-vlaue p1 for hypothesis test
# Ho: B1 = 0 versus H1: B1 != 0
t.test(genetic_data$y,genetic_data$X14046)
```
p-value = 2.2e-16

Using alpha = .05, we reject the null hypothesis and can conclude that gene1 is statistically significant.

Q3 B) Repeat using genes 2-10 as the predictor
```{r}

genetic.lm.genes <- lm(y~X15850+X23006+X22110+X29566+X21092+X30116+X22029+X15224+X25141, data=genetic_data)

summary(genetic.lm.genes)
coefficients(summary(genetic.lm.genes))

```
The lower the p-value -> the greater the statistical significance of the observed difference

=> Gene6 (X21092) has the most significant association with the syndrome


Q4: A) Plot the data using the pairs(stackloss) function
```{r}
data(stackloss)
names(stackloss)

# plot data 
pairs(stackloss)
```

B) Fit a multiple regression model 
```{r}
model <- lm(stack.loss ~ Air.Flow + Water.Temp + Acid.Conc., data=stackloss)
summary(model)

```

Q4 C): Construct 90% confidence intervals for the coefficients
```{r}
confint(model, level=.90)
```
The 90% confidence intervals for each coefficient

Q4 D) Construct a 99% percent prediction interval for a new observation when Airflow = 58, Water.temp = 20, ACid = 86
```{r}
newdata = data.frame(Air.Flow=58, Water.Temp=20, Acid.Conc.=86)
## predict with new frame
predict(model, newdata, interval="predict", level=.99)
```

The 99% prediction interval of stackloss when Air Flow=58, Water Temperature = 20, and Acid Concentration=86 is between 4.759959 and 24.06133

Q4 E) Test Null hypothesis Ho: B3 = 0, What is the p-value. What is the conclusion at alpha = .10

Alternate hypothesis: H1: B3 != 0


```{r}
coefficients(summary(model))

```


The p-value is .3440461, using an alpha = .10, we can conclude that the null hypothesis is true. There is no statistical significance between stackloss and acid concentration.



